Cycle Route Uptake and Scenario Estimation (CRUSE): an approach for developing strategic cycle network planning tools

### 2024-03-26 paper.qmd
add Declan comments

Practitioners in Kildare County Council that the evidence generated *was already valuable* the development of new cycle routes, especially when making the case during planning applications.
They emphasized that cycle networks are often fragmented, and the tool allows them to demonstrate how individual local schemes can contribute to a larger, interconnected network.
An unexpected piece of feedback from the workshop was the preference for visualizations that directly compare the fast and quiet routes, as they recognize the importance of both types of routes.
In response to this we provided the 'Route types' page for each county, although this functionality is not available in the landing page map due to other feedback requesting a simpler interface.
In comparison with another tool (the National Transport Authority's cycling propensity tool), the CRUSE tool was seen by practitioners in Kildare as more user-friendly, with the ability to generate evidence at the route level, which is crucial for planning new cycling infrastructure.
Overall, the workshop participants concluded that the approach filled gap in evidence and will greatly assist in the development of new cycling projects.

@@

In a subsequent interview with a practitioner using the evidence generated by the CRUSE approach in their everyday work, we received the following feedback.
The outputs were a "great tool for getting started with project outline documents" and "ideal for generating a strategic rationale for estimating demand and building the business case in the absence of cycle count data" (Declan Keenan, Senior Transport Planner working on active travel, personal communication).
Declan also noted the alignment between local knowledge, cycle count data, and the outputs of the approach: "It's a good representation of what's going on" and the networks allign well in terms of network "shape" and the proportion of trips of trips on different parts of the network.

Findings from the qualitative results also suggest changes which have not been implemented due to time constraints, including the provision of route netework data showing only "gold standard" POWSCAR data (we did make this data available in the data downloads page), the need for further validation of results (especially in rural areas where counter datasets are scarce) and the need for training on the tool to ensure that it is well-understood and used effectively.
These and other potential areas for further research and development are discussed in the next section.

>>>>

Practitioners in Kildare County Council found that the evidence generated *was already valuable* for the development of new cycle routes, especially when making the case during planning applications.
They emphasized that cycle networks are often fragmented, and the tool allows them to demonstrate how individual local schemes can contribute to a larger, interconnected network.
An unexpected piece of feedback from the workshop was the preference for visualizations that directly compare the fast and quiet routes, as they recognize the importance of both types of routes.
In response to this we provided the 'Route types' page for each county, although this functionality is not available in the landing page map due to other feedback requesting a simpler interface.
In comparison with another tool cycling propensity tool, the CRUSE tool was seen by practitioners in Kildare as more user-friendly, with the ability to generate evidence at the route level, which is crucial for planning new cycling infrastructure.
Overall, the workshop participants concluded that the approach filled gap in evidence and will greatly assist in the development of new cycling projects.

@@

In a subsequent interview with a practitioner using the evidence generated by the CRUSE approach in their everyday work, we received the following feedback.
The outputs were found to be a "useful tool for getting schemes started; and provides good detail to supplement the strategic rationale for intervention within Project Outline Documents, via the cycle friendliness maps and the baseline / near market cyclist estimates”. As a result, the CRUSE tool “is ideal for generating a strategic rationale for estimating demand and building the business case in the absence of cycle count data; it is also very useful at the detailed design stage and the CRUSE estimates can (and have been used) to supplement the economic appraisal of cycling schemes” (Declan Keenan, Senior Transport Planner TII working on active travel, personal communication).
Declan also noted a strong alignment in the South Dublin area between cycle count data, and the outputs of the approach: “It’s forms a strong representation of what’s going on in the areas where TII have collated some cycling count data, e.g. in the vicinity of the M50 in the areas of Ballinteer and Sandyford. Some additional network checking against count data would also be useful, to ensure good representation across other settlements.” Where cycling data is available, the networks align well in terms of network “shape” and the proportion of trips on different parts of the network.

Where cycling data is available, the networks align well in terms of network “shape” and the proportion of trips on different parts of the network. There were two suggestions from Declan which have not yet been implemented: the provision of data showing POWSCAR and recreational estimates/data separately, and the need for training on the tool to ensure that all staff can use it effectively. These, and other potential improvements, are discussed in the next section.
